Corridor Project Proposal

### 2024-01-26 corridor_docs/proposal/eec_corridor_proposal.Rmd
adding plots for each experiment

1. Prediction 1: Species Richness, Species Diversity, and Functional Diversity will be higher in patches connected by corridors than in unconnected patches

<!-- 2. Predictions 2: Corridors facilitate inter-patch movement of dung beetle species,  -->
2. Prediction 2: Corridors are suitable habitat for all species. However, the speed at which in individuals move through corridors and the probability of successful inter-patch movement is size-dependent (i.e., larger beetle species move more quickly and are more likely to reach the connected patch).

3. Dung removal rates will be highest in connected patches. This is due to the higher functional diversity of beetles in these locations.

@@

### Methods

@@

1. Design Prediction 1 (pitfall sampling)
1. Design Prediction 2 (MRR)
1. Design Prediction 3 (Dung Removal Experiment)

![Sampling Design](./images/srs_wt.png){fig-align="center" width=95% height=95%}

>>>>

1. **Prediction 1:** Species Richness, Species Diversity, and Functional Diversity will be higher in patches connected by corridors than in unconnected patches

<!-- 2. Predictions 2: Corridors facilitate inter-patch movement of dung beetle species,  -->
2. **Prediction 2:** Corridors are suitable habitat for all species. However, the speed at which in individuals move through corridors and the probability of successful inter-patch movement is size-dependent (i.e., larger beetle species move more quickly and are more likely to reach the connected patch).

3. **Prediction 3:** Dung removal rates will be highest in connected patches. This is due to the higher functional diversity of beetles in these locations.

@@

### Study Site and Methods

@@

_Methods_
**Methods - Prediction 1:**

To calculate species richness, species diversity, and functional diversity we will sample the dung beetle community in each location with pitfall traps baited with cow dung (Spector 2006, Andrade et al. 2011). In each site we will arrange traps in a grid, with traps separated from each other by 25 m. The traps will be 25 m from the edge of the patch to avoid attracting beetles from the matrix outside of the patch. Traps in each location will be set on the same day and left out for 24 hours.

The traps, modified from the design of Wood and Kaufman (2008), are constructed from 6 inch sections of (4?) inch(cm) diameter PVC pipe fitted with a screen on the bottom and a -- cm diameter funnel on top with an extended --cm plastic platform used to suspend the trap over a hole in the ground. The traps will be set level to the ground by placing them into a semi-permanent hole that was maintained by a larger section of PVC while traps were not in use. Dung is wrapped in paper towel to exclude beetles and suspended over the trap on a wire cloth platform. Note that although the traps of Wood and Kaufman (2008) were designed as a live catch trap, ours will be converted to a wet trap by placing a --ml cup containing 95% ethanol inside the trap (the original design allows some beetles to escape and for predatory beetles to enter the traps).

The dung used to bait traps will be collected from the University of Florida’s Beef Teaching Unit from cattle that were unsupplemented – (with what). It is  collected fresh in the morning, homogenized, and formed into approximately (_ml) balls and frozen at (_degrees Celsius) until deployment. While some dung beetles have a clear preference for a particular type of dung, most species are generalists that use a broad diversity of vertebrate excrement (cite - check Gordon 1983, and Mart´in Piera and Lobo 1996, and see what Woodruff has to say; see page 13 of Haffter and Edmonds). All beetles captured will stored in 95% ethanol and identified to species or morphospecies using Woodruff (1973). A collection of voucher specimens will be deposited at the FLMNH and ------.

1. Design Prediction 1 (pitfall sampling)

- 5 traps in a grid in the central 50 x 50 m of each patch + 1 sampling grid in the matrix. The 25 m buffer from the edge is to avoid beetles coming into the plot from the matrix.
- will try to test buffer width with marked beetles released at different distances, including from outside matrix
- 4 arrays per blocks = 36 traps per array, 7 blocks = 252 traps total
- need to make sure - ebnough "coverage" without a buffer too small, meaning bugs from matrix come into patch

![Sampling Design](./images/plots1.png){fig-align="center" width=50% height=50%}

**Methods - Prediction 2:** 
2. Design Prediction 2 (Dispersal)

- using connected patches + unconected: 2 dung baits in center patch. one at ende closest to isolated patch, one at end of corridor,. release marked beetles at ende of isolated patch and entry of corridor. Check at intervals of X to see if they will make it.
- to consider: passive traps to see if they are moving through? shouidl we start experiment with smaller distances into the orridor and then eventually the full length? 
- tech to follow them?

![Sampling Design](./images/plots2.png){fig-align="center" width=50% height=50%}

**Field Methods - Prediction 3:** 
3. Design Prediction 3 (Dung Removal Experiment)

- field experiment: 300 g of dung on surface, 300 g on surface covered with mesh (controls for dessication), weigh after 48 hours
-  to consider: gather and sieve soil from outside, place in shallow pot, and place dung on that. Allows breakdown of soil without getting into their soils, if this is a concern
- repeat biweekly for X months 
- lab experiment: meso-cosms in GNV based on results from diversity sampling 
- need to make sure each point inside a plot is independent, and that each sampling date is independent

![Sampling Design](./images/plots3.png){fig-align="center" width=50% height=50%}
